### 2016-07-06 image_paper3/text_files/image_texts/col3byall.tex
adding prelude to into. section

\begin{figure*}
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and11.png}
        \label{fig: col3and11_dist}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and12.png}
        \label{fig: col3and12_dist}
    \end{subfigure}
        \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and13.png}
        \label{fig: col3and13_dist}
    \end{subfigure}
        \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and14.png}
        \label{fig: col3and14_dist}
    \end{subfigure}
        \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and15.png}
        \label{fig: col3and15_dist}
    \end{subfigure}
        \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and16.png}
        \label{fig: col3and16_dist}
    \end{subfigure}
        \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and17.png}
        \label{fig: col3and17_dist}
    \end{subfigure}
        \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and18.png}
        \label{fig: col3and18_dist}
    \end{subfigure}
        \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and19.png}
        \label{fig: col3and19_dist}
    \end{subfigure}
        \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and20.png}
        \label{fig: col3and20_dist}
    \end{subfigure}        \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and21.png}
        \label{fig: col3and21_dist}
    \end{subfigure}
            \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and22.png}
        \label{fig: col3and22_dist}
    \end{subfigure}
            \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and23.png}
        \label{fig: col3and23_dist}
    \end{subfigure}
            \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and24.png}
        \label{fig: col3and24_dist}
    \end{subfigure}
            \hfill
    \begin{subfigure}[b]{0.25\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and25.png}
        \label{fig: col3and25_dist}
    \end{subfigure}
    \caption{Results from increasing dimension of input data from 11 to 25}
    \label{fig: inc_D_col3s}
\end{figure*}