### 2016-07-06 image_paper3/text_files/image_texts/subset9.tex

\begin{figure}
 \begin{subfigure}[b]{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/image_subsets/subset9_dist_with_hits_t.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.5\textwidth}
    \includegraphics[width=\textwidth]{../../images0.01/M31/2D/image_subsets/subset9_dist_with_hits_v.png}
    \end{subfigure}
    \caption{Results from subset 9}
    \label{fig: subset9}
\end{figure}