### 2016-07-17 image_paper3/text_files/image_texts/PAHS_or_not_PAHs.tex
fig.s and table corrections

\begin{figure}
    \begin{subfigure}[b]{0.8\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols11and26.png}
    \caption{Without PAHs}
    \label{fig: wt_pahs}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.8\textwidth}
    \centering
        \includegraphics[width=\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and10.png}
    \caption{Only PAHs}
    \label{fig: only_pahs}
    \end{subfigure}
    \caption[Two dimensional self-organizing map of M31 for no PAHs and only PAHs subsets]{Similar to Fig.~\ref{fig: all_derived_ones}, here we show the result of a) SOM from data excluding PAHs; b) SOM from PAHs data.}
    \label{fig: PAHS_or_not_PAHs}
\end{figure}