### 2016-07-17 image_paper3/text_files/image_texts/all.tex

\begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{../../images0.01/M31/2D/diff_dimension/combine_2D_data_between_cols3and26.png}
    \caption[Two dimensional self-organizing map of M31 data]{Two dimensional SOM of M31 using $10\times10$ grid. Similar to Figs.\ref{fig: sample} and \ref{fig: M31_nets_1d}, the hexagonal shapes show the neurons and the grey cycle colours between neurons show the differences between weight of each neuron.}
    \label{fig: all_derived_ones}
\end{figure}